---
title: "STA 5820 Chapter 3 lab: Linear Regression"
author: "魏上傑"
date: "2023-04-22"

output:
  pdf_document:
    latex_engine: xelatex


  
mainfont: "Times New Roman"  # 指定英文字型
fontsize: "12pt"
papersize: "a4"
geometry: "margin=1.5cm"
lang: "zh-tw"
documentclass: ctexart

---

Some supplemental codes for linear models.

# Boston Data set

```{r}
library(MASS) # Boston data
head(Boston) # median value of homes by town
```

```{r}
# see help for definition of variables
attach(Boston) 
```

# Scatter plot for crime rate vs. median house value

```{r}
plot(crim, medv)
```

```{r}
hist(crim, breaks = seq(0, 90, 5))
```

```{r}
hist(medv, breaks = seq(0,50,5))
```

The crime rate is a skewed right, and needs a concave transformation to make the relationship between the two variables more linear.

```{r}
plot(log(crim), medv)
```

While the median house values seem to be truncated at 50 ($50,000), the crime rate and the median house value has roughly a linear relationship.

# Fitting a linear model

```{r}
LM1 <- lm(medv~ log(crim))
summary(LM1)
```

$medv =-1.93 log(crim) +21 +\epsilon$

$e^{medv}=crim\times e^{-1.93}$

```{r}
# confidence band, prediction band
confint(LM1, level = 0.99) # 99% confidence interval (default=95%)
```

```{r}
CI <- predict(LM1, interval="confidence") # confidence interval
head(CI)
```

```{r}
PI <- predict(LM1, interval = "prediction") #prediction interval
head(PI)
```

```{r}
ORD <- order(crim) # index for crim in increasing order
head(ORD)
```

```{r}
plot(log(crim), medv, ylim = c(-10, 60))
matlines(log(crim)[ORD], CI[ORD,], type="l", col=c(1,2,2), lty=c(1,2,2))
matlines(log(crim)[ORD], PI[ORD,], type="l", col=c(1,3,3), lty=c(1,2,2))
```

Roughly 95% of the observations are within the prediction bands (green). As there are many (506) observations, the confidence band (red) is narrow, indicating that the estimated regression line (black) is reasonably accurate.

# leverage

```{r}
# plot and leverage
hist(hatvalues(LM1))
```

```{r}
sort(hatvalues(LM1))[501:506] # 6 largest hat values

```
There are no extreme hatvalues, compared to the others.


```{r}
which.max(hatvalues(LM1)) # point with largest leverage
```

```{r}
MAX3 <- order(hatvalues(LM1))[504:506] # points with 3 largest leverages
MAX3
```



```{r}
library(car) # for pointLabel 
```

```{r}
plot(log(crim), medv)
pointLabel(x=log(crim)[MAX3], y=medv[MAX3], labels = as.character(MAX3))
```

```{r}
plot(dis, medv)
```

```{r}
plot(nox, medv)
```

```{r}
# ANOVA (Analysis of Variance)
LM2 <- lm(medv~ log(crim)+dis+nox)
summary(LM2)
```

```{r}
anova(LM2)
```

```{r}
LM2b <- lm(medv~ log(crim)+dis+nox) 
# change of order doesn't matter here
summary(LM2b)
```

```{r}
anova(LM2b) # change of order matters in ANOVA
```

```{r}
anova(LM1, LM2) # see if LM2 is significantly better than LM1
```

