---
jupyter:
  jupytext:
    formats: ipynb,Rmd,R
    text_representation:
      extension: .Rmd
      format_name: rmarkdown
      format_version: '1.2'
      jupytext_version: 1.9.1
  kernelspec:
    display_name: R
    language: R
    name: ir
output: pdf_document
---

# Lab 1

This is loosely based on notes by Dr. Leemis: [here](http://www.math.wm.edu/~leemis/r.pdf). If you're not familiar with `r` I highly recommend you go through this. 

## Notebook basics

Code notebooks allow interweaving of 

1. rich-text input (html, markdown, etc.)
2. code input
3. code output

```{r}
print('hello darkness my old friend')
```



We can write notebooks using a couple different IDEs, popular being 

1. [jupyter](https://jupyter.org/install) which is a bit more advanced and requires installing [anaconda](https://www.anaconda.com/)

2. [rstudio](https://rstudio.com/products/rstudio/)

either can be used to write code in [R](https://www.r-project.org/)


## R as a calculator

We can use R as a calculator

```{r}
1+1
```

```{r}
1+2*5
```

there are also some built-in constants in R

```{r}
pi
exp(1)
```

there are also special infinite values

```{r}
1/0
-1/0
0/0
```

to denote nothing we have

```{r}
NULL
```

to denote a missing value

```{r}
NA
```


# Variable assignment and simple objects

One can assign variables with either an `=` or `<-`

```{r}
x = 1
x
```

```{r}
x <- 1
x
```

## Vectors

vectors are made with the `c` command

```{r}
x = c(5, 3, 7)
x
```

we get the elements with `[]`

```{r}
x[1]
```

negative numbers selects all but those elements


```{r}
x[c(-1,-2)]
```


we can make a vector out of any type

```{r}
truths <- c(TRUE, FALSE, TRUE)
truths
```

if we subset by boolean then it select those elements with `TRUE`

```{r}
x
```

```{r}
truths
```

```{r}
x[truths]
```

we can make consecutive integers using:

```{r}
1:10
```

or the `seq` function

```{r}
seq(1, 20, by=3)
```

## matrices

we can make a matrix using the ``matrix`` function

```{r}
X <- matrix(1:25,nrow=5,byrow=TRUE)
X
```
or the `array` function

```{r}
Y <- array(25:49, c(5,5))
Y
```
matrix multiplication is done with the operator ``%*%``

```{r}
X %*% Y
```

I can extract or assign individual elements with `[,]`

```{r}
X[1,2]
```

```{r}
X[1,2] <- 3
X
```

or entire rows/columns as follows

```{r}
X[,1]
```

```{r}
X[1,]
```


# Flow Control

if statements are as follows

```{r}
A = 5
if(A==0){
    print("A=0")
} else if(A == 1){
    print("A=1")
} else if(A==2) {
    print("A=2")
} else {
    print("A is not 0, 1 or 2")
}
```

we can also make `for` loops

```{r}
for(i in 1:10){
  print(i)
}
```

and `while` loops

```{r}
i=1
while(i <= 10){
  print(i)
  i = i+1
}
```

# major types

numeric

```{r}
c(1,2,3)
```

```{r}
class(c(1,2,3))
```

strings

```{r}
c('hello','darkness','my','old','friend')
```


```{r}
class(c('hello','darkness','my','old','friend'))
```

boolean

```{r}
c(TRUE,FALSE)
```

```{r}
class(c(TRUE,FALSE))
```

factors for discrete variables

```{r}
fctr = as.factor(c("hello", "darkness"))
```

```{r}
fctr
```

```{r}
class(fctr)
```

```{r}
as.numeric(fctr)
```

```{r}
fctr*2
```

# data-frames and matrices:

matrcies all have to be the same type:

```{r}
X = array(0, c(5,5))
X
```

```{r}
X = array(LETTERS[1:25], c(5, 5))
X
```

so if I have a character `X` and I assign the number `5` to the first column, what happens?

```{r}
X[,1] = rep(5,5)
X
```
```{r}
X[,1]
```

this is not the number `5`, its the character `'5'`

The proper way to have mixed types is to have a `data.frame`

```{r}
df = data.frame(x=rep(5,5),y=LETTERS[1:5])
df
```

I can access the elements like matrices

```{r}
df[1,1]
df[1,2]
```

or with names

```{r}
df$x
df$y
```

```{r}
df[["x"]]
df[["y"]]
```

